Synthesizing Quotient Lenses

### 2018-03-11 papers/QuotientLenses/main.tex
minor fixes and proposed alternate qre example

\emph{Bijective lenses} are amongst the most restrictive. 
A bijective lens $\ell : S \Leftrightarrow T$ specifies a bijection from
a source set $S$ to a target set $T$:
\begin{equation}\label{bijectivelenslaws}
\ell.\get \; (\ell.\lput \; t) = t \text{, and } \ell.\lput \; (\ell.\get \; s)
= s
\end{equation}
Since these lenses demand that the \emph{get} and \emph{put} functions form a bijection, 
no information is lost during translation from one data format to another. 
While this very strong guarantee can be useful, more often
some of the information in one or both of the data formats is
superfluous. For instance, the number of white space characters
between data items, the capitalization of various strings, the
sequence of fields in record, or the order of items in a list are
often irrelevant.  Many
data formats have their own ad hoc rules as well. In \bibtex{}, for example,
an author can be specified as
``\cd{lastname, firstname}'' or as ``\cd{firstname lastname}''---
the specific choice does not matter.

{\em Quotient lenses}~\cite{quotientlenses} loosen the restrictions of
bijective string lenses by including equivalence
relations on the source and target formats so that the round-tripping lens laws
hold modulo these equivalence rather than modulo equality. Quotient lenses thus
make it possible to synchronize many more data formats than would otherwise be
possible by ignoring details deemed inessential.

@@

Our solution to this shortcoming is to introduce {\em Quotient Regular
Expressions} (QREs), which enable programmers to simultaneously specify a
regular expression and a set of canonical representatives for matching
strings.  Given an QRE, we can automatically infer the corresponding
canonization functions.  
\ksf{Define comma\_name here as a really simple example? }

@@

Another QRE primitive is the sequential composition combinator,
written ``;''.
For an example of its use, suppose we have already defined a QRE
\codefont{comma_to_space} that accepts names in either space-separated
or name-separated format and chooses the space-separated format as canonical.
Suppose we have also already defined a
QRE \codefont{project_in_space}, which projects away extra white space
between the first and last name in the space-separated name format. 
We can use the sequential composition combinator to combine these two QREs:
\begin{lstlisting}
let name = project_in_space ; comma_to_space
\end{lstlisting}

@@

(* squash QRE maps comma_name to space_name *)
let comma_to_space = squash comma_name $\to$ space_name using (get comma_to_space)

@@

(* Project away whitespace in space_name representation *)
let project_in_space = first_name . p_wsp . last_name

(* Compose the two QREs to get a canonical representation for a name *)
let name = project_in_space ; comma_to_space

(* define rest of bibtex fields *)
let name_sep = p_wsp . "and" . p_wsp
let bib_names = name | (name . name_sep . (name . name_sep)$^*$ . name)
let bib_author = "author = \"" . bib_names . "\""
let title = word | (word . p_wsp . (word . p_wsp)$^*$ . word)
let bib_title = "title = {" . title . "}"

(* allow any permutation of fields interspersed with arbitrary whitespace *)
let bib_fields = perm (label, bib_author, bib_title) with (project ("," . wsp+) $\to$ ",\n")
let bibtex = "@book{" . bib_fields . "}"
\end{lstlisting}
\caption{QRE specification of \bibtex{} records. \ksf{Is all this code
correct?  Does it run through the system correctly?  I'm suspicious of
the comma\_to\_space definition which seems to be recurisive.  If it's
correct, we should comment on it.}\ksf{In what sense is this a QRE
spec vs a QRE lens spec?}}
\label{fig:example-qre}
\end{figure}

>>>>

\emph{Bijective lenses} are amongst the most restrictive. 
A bijective lens $\ell : S \Leftrightarrow T$ specifies a bijection from
a source set $S$ to a target set $T$:
\begin{equation}\label{bijectivelenslaws}
\ell.\get \; (\ell.\lput \; t) = t \text{, and } \ell.\lput \; (\ell.\get \; s)
= s
\end{equation}
Since these lenses demand that the \emph{get} and \emph{put} functions form a bijection, 
no information is lost during translation from one data format to another. 
While this very strong guarantee can be useful, more often
some of the information in one or both of the data formats is
superfluous. For instance, the number of white space characters
between data items, the capitalization of various strings, the
sequence of fields in a record, or the order of items in a list are
often irrelevant.  Many
data formats have their own ad hoc rules as well. In \bibtex{}, for example,
an author can be specified as
``\cd{lastname, firstname}'' or as ``\cd{firstname lastname}''---
the specific choice does not matter.

{\em Quotient lenses}~\cite{quotientlenses} loosen the restrictions of
bijective string lenses by including equivalence
relations on the source and target formats so that the round-tripping lens laws
hold modulo these equivalences rather than modulo equality. Quotient lenses thus
make it possible to synchronize many more data formats than would otherwise be
possible by ignoring details deemed inessential.

@@

Our solution to this shortcoming is to introduce {\em Quotient Regular
Expressions} (QREs), which enable programmers to simultaneously specify a
regular expression and a set of canonical representatives for matching
strings.  Given an QRE, we can automatically infer the corresponding
canonization functions.  
\ksf{Define comma\_name here as a really simple example? }
\afm{Right now comma\_name is just a regular expression.  Further details in Fig
2}

@@

Another QRE primitive is the sequential composition combinator,
written ``;''.
For an example of its use, suppose we have already defined a QRE
\codefont{comma_to_space} that accepts names in either space-separated
or name-separated format and chooses the space-separated format as canonical.
Suppose we have also already defined a
QRE \codefont{project_in_space}, which projects away extra white space
between the first and last name in the space-separated name format. 
We can use the sequential composition combinator to combine these two QREs:
\begin{lstlisting}
let name = comma_to_space ; space_name_p
\end{lstlisting}

@@

(* squash QRE maps comma_name to space_name *)
let comma_to_space = squash comma_name $\to$ space_name using c_to_s

@@

(* Project away whitespace in space_name representation *)
let space_name_p = first_name . p_wsp . last_name

(* Compose the two QREs to get a canonical representation for a name *)
let name = comma_to_space ; space_name_p

(* define rest of bibtex fields *)
let name_sep = p_wsp . "and" . p_wsp
let bib_names = name . (name_sep . name)$^*$)
let bib_author = "author = \"" . bib_names . "\""
let title = word . (p_wsp . word)$^*$
let bib_title = "title = {" . title . "}"

(* allow any permutation of fields interspersed with arbitrary whitespace *)
let bib_fields = perm (label, bib_author, bib_title) with (project ("," . wsp+) $\to$ ",\n")
let bibtex = "@book{" . bib_fields . "}"
\end{lstlisting}
\caption{QRE specification of \bibtex{} records.\ksf{In what sense is this a QRE
spec vs a QRE lens spec?}\afm{I don't really like the way this is set up.  The
use of the sequential composition operator makes the QREs appear really
operational, and in my opinion makes them seem less cool, and distinct from
canonizers, than they actually are.  I include a proposed alternative in the
next figure.  However, this would mean that the previous section which describes
the sequential composition operator would require it be talked about without the
concrete example.}}
\label{fig:example-qre}
\end{figure}

\begin{figure}
\begin{lstlisting}
let wsp = [ \n\t\r]+
let wsp_s = project wsp $\to$ " "

(* define name representations with a space and with a comma *)
let space_name = first_name . wsp_s . last_name
let comma_name = last_name . "," . wsp_s . first_name

(* squash QRE maps comma_name to space_name *)
let name = squash comma_name $\to$ space_name using c_to_s

(* define rest of bibtex fields *)
let bib_names = name . (wsp_s . "and" . wsp_s . name)$^*$
let bib_author = "author = \"" . bib_names . "\""
let title = word . (wsp_s . word)$^*$
let bib_title = "title = {" . title . "}"

(* allow any permutation of fields interspersed with arbitrary whitespace *)
let bib_fields = perm (label, bib_author, bib_title) with (project ("," . wsp) $\to$ ",\n")
let bibtex = "@book{" . bib_fields . "}"
\end{lstlisting}
\caption{QRE specification of \bibtex{} records.\ksf{In what sense is this a QRE
spec vs a QRE lens spec?}}
\label{fig:example-qre-anders-proposal}
\end{figure}
